# POE-group-proj

```{r}
install.packages("ggplot2")
```

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data <- Group_7_data

ggplot(data,
       mapping = aes(x = `Area of park (km2)`,
                     y = `Number of species`))+
  theme_bw() +
  geom_point(size = 1, alpha = 0.5) + 
  geom_smooth(method = lm, se = TRUE)+
  labs(title = "Scatter Plot of the number of bird species against the area of parks",
         subtitle = "2024", 
         x = "Area of park, in km2",
         y = "Species richness",
         colour = " ") 



```
square model? ANOVA model? 

```{r}
data %>%
ggplot(mapping = aes(x = `Area of park (km2)`)) +
    geom_histogram(binwidth = 0.0001, fill = "lightblue", color = "black") 
```
the histogram is skewed to left too much
maybe square the area of park and analyse the distribution based on it? 
That way, small values are more after squared, and the normality of data is increased? 

```{r}
data %>%
  mutate(squared_area = (`Area of park (km2)`)^2) %>%
  ggplot(mapping = aes(x = `squared_area`)) +
    geom_histogram(binwidth = 0.0005, fill = "lightblue", color = "black")

```
```{r}

ANOVA_data <- data  %>%
  mutate(size = case_when(
    `Area of park (km2)` <= 0.01 ~ "Small", 
    `Area of park (km2)` > 0.01 & `Area of park (km2)` <= 0.17 ~ "Medium", 
    TRUE ~ "Large"
  )) %>%
  reframe(`Area of park (km2)`, `Number of species`, size)

mean <- ANOVA_data%>%
  group_by (size) %>% 
  summarise (mean_area =  mean (`Area of park (km2)`) ,
             mean_richness = mean(`Number of species`),
             stdev = sd (`Number of species`), 
             se = sd (`Number of species`)/ sqrt (n () ))

mean_data <- ANOVA_data %>%
  group_by(size) %>%
  summarise(mean_richness = mean(`Number of species`, na.rm = TRUE))



```
Preparation for ANOVA, devided parks by three different types of size, small, medium and large. 
The boundaries of each park size were determined by the histogram of park area above. 

summarised statistical values

```{r}
z <- ggplot(data = ANOVA_data, 
       mapping = aes(x = size,
                     y = `Number of species`),
                     colour = size) +
  geom_point(size = 2, alpha = 0.8, shape = 21) +
  geom_point(stat = "summary", fun="mean", size = 3)


z <- z+ geom_errorbar (data = mean, 
                 aes(x=mean$size, 
                     ymin = (mean$mean_richness - mean$se),
                     ymax = (mean$mean_richness + mean$se), col=size),
		width=0.1) +
    theme_bw()


p <- plot( mean_richness ~ as.factor(size), data = mean, ylim=c(0,20), 
        xlab= "Size of Park", ylab = "Species richness", col = "black", cex = 1) + 
  points(`Number of species` ~ as.factor(size), data = ANOVA_data, 
       col = as.factor(ANOVA_data$size), cex = 0.7) +
  arrows(x0=1:3, 
	y0=mean$mean_richness - mean$se, 	x1=1:3, 
	y1=mean$mean_richness + mean$se,
	code=3, length=0.2, angle=90, col= "orange", lwd=1)

```

```{r}


```




